Depression and suicidality as costly signals of need

### 2021-04-15 Paper.rmd
Final changes before turning in draft to Anne

Major depression and suicide are generally viewed as pathological responses to stress that elicit aversive responses from others. Alternatively, depression and suicide attempts might be the result of adaptations for reliably signaling need and bargaining for increased support from reluctant others. To address this hypothesis, we examined responses to emotional signals in an experimental vignette study involving claims of substantial need in the presence of conflicts of interest and private information about whether the signaler was actually in need. In our sample of 1,240 participants recruited from Amazon Mechanical Turk, costlier signals like depression and suicide attempts resulted in greater perceptions of need and increased likelihood of support compared to simple verbal requests and crying without further symptoms. Additionally, the effect of signal type on likelihood of support was largely mediated by the effect of signaling on belief in need as predicted. Our results suggest that depression and suicide attempts, apparent human universals, are a relatively successful way of signaling honest need and eliciting support, consistent with costly signaling explanations.

@@

Durkheim’s pioneering study of suicide [@durkheim1897] helped launch sociology as a quantitative discipline. Sociological investigations of suicide flourished through much of the 20th century, but were supplanted by medical and public health approaches as the biomedical model of mental health took hold in the late 1970's and early 1980's [@wray2011; @syme2020]. The biomedical model tends to frame suicide as a problem with the individual, e.g., as one involving psychiatric disturbance, drug addiction, or stress sensitivity [@wray2011]. This theoretical approach has made little progress: Rates of suicide, which vary widely by country, have remained relatively flat in countries for which there are long-term data, such as the US, UK, Japan, and Sweden [@owidsuicide].

@@

Evolutionary theories of depression tend to either ignore suicidality, or take it as evidence that depression is dysfunctional. Wakefield, for instance, who has devoted much of his career to arguing that most cases of depression are simply severe but normal sadness and grief, nevertheless takes suicidality to indicate pathology for some cases of depression [e.g., @wakefield2017]. @nettle2004 argues that the costs of depression, including suicidality, outweigh any benefits, and therefore adaptationist models of depression are implausible. @gilbert2006 sees suicidality as an example of a dysregulated escape defense [for reviews of evolutionary theories of depression, see @].

and anthropological explanations for low mood have become more common. In general, these approaches see sadness and minor depression as ways to motivate adaptive behavior changes in response to negative outcomes in one's environment (cite Nesse and others), mainly differing in how they view major depression. While some approaches present most forms of major depression as dysfunctional (inflammation and some of Nesse), others have extended the logic on the benefits of low mood and reduced behavior to more severe instances of sadness. Of the possible benefits, the ability of depression to help solve fitness-relevant in times of social conflict is the most emphasized, with depression being suggested to help by reallocating resources with the body for problem solving (cite ARH), reducing the costs of interacting with higher status individuals (cite), or serving as a means to both signal one's need and bargain for increased support (cite).

As with suicide, research on depression since the 1980's has been dominated by biomedical approaches of limited success [@patten2016; @ormel2019].

and there is still substantial uncertainty about why both are as common as they are (cite). Likely due to  the suffering they cause, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others).

@@

Likely the most reliable finding regarding depression and suicidality is that most cases are closely tied to major negative life events [@Hammen2005; @Mazure1998; @devries2011; @ellsberg2008]. Compared to non-depressed individuals, those with depression tend to report at least twice as many instances of negative events than non-depressed individuals [@Mazure1998] and more negative events than those with schizophrenia and bipolar depression across multiple studies [@Paykel1994]. In longitudinal studies, the onset of depression has been reported to quickly follow the negative event [@KendlerEtAl1999] or coincide with periods where conflict is likely to increase prior to it (e.g., depression starting before a divorce rather than after; @rosenstrom2017). 
<!-- look for other longitudinal studies -->

This strong connection between negative life events and depression along with the adversity and distress that likely accompanies these events has led to the suggestion that this relationship is causal [@KendlerEtAl1999]. Support for this comes from the fact that the relationship between negative events and depression holds true when looking solely at events outside of one's control, showing that the connection is unlikely to be driven solely by individuals who are already depressed creating situations where negative events are likely to be common [@Hammen2005; @KendlerEtAl1999]. In addition to this, twin studies have shown that one's history of negative events is still a strong predictor of major depression when controlling for genetic similarity and that part of the heritability of depression stems from the heritability of negative events, like divorce and family conflict [@KendlerBaker2007; @KendlerEtAl1999].

Of the types of negative life events associated with depression, those which involve conflict tend to be more strongly associated with depression [@Hammen2005; @Mazure1998]. Marital problems, bullying, and the existence of abusive relationships are all commonly cited risk factors for depression [@KendlerEtAl1995; @KendlerEtAl1999; @klomek2007; @klomek2019], with  assault (both sexual and non-sexual), in particular, greatly increasing one's risk of depression [@KendlerEtAl1995; @KendlerEtAl1999].  Other notable risk factors like loss of a loved one, financial problems, job loss, or severe or prolonged illness, might seem less related to conflict. However, the ability for others to provide meaningful support to limit the harm caused in times of need [@SugiyamaSugiyama2003] combined with the ubiquity of conflicts of interests mean that many types of problems will eventually involve social conflict even if they did not start that way [@Hagen2003]. In line with this reasoning, conflict, as a whole, was found to be associated with depression among the Tsimane, particularly when it involved non-kin [@StieglitzEtAl2015a].

As depression is the leading cause of suicide compared to mental disorders [@who2001], it should not be surprising that suicidality also often coincides with social adversity and conflict. Grief over the loss of a loved one [@firth1936: @Firth:1961; @devries2011], bullying [@alavi2017; @butler2019; @ford2017; @wang2020], and instances of social transgression [@Firth:1961; @Malinowski:1932; @SymeHagen2018; @lester1997] have all been reported to lead to suicidal behavior in both industrialized and traditional societies. Of the recorded instances of suicide attempts in the Human Relations Area Files (HRAF), 77.4% of attempts were found to be linked to times of social conflict, 88.7% were linked to fitness threats in general, and 84.9% were linked to situations in which attempters were powerless to resolve their situation alone [@SymeEtAl2016]. As with depression, assault is an particularly powerful risk factor for suicidal behavior based on multiple large-scale multi-country studies [@devries2011; @ellsberg2008] and more localized examinations [@reigstad2017]. 
<!-- need to find something more recent citation than who2001 -->
## Depression and suicide attempts as both signals and bargaining tools

@@

Due to the ease with which signals can be faked [@Zahavi1975; @mokkonen2016] and the costs of accepting a deceptive signal as reliable [@dawkins1978; @johnstone1993; @Hasson1994], demonstrating that one’s need is genuine has likely been a recurrent adaptive problem since humans evolved the capacity to provide instrumental support to one another. One possible solution to this is to produce signals that are unprofitable to fake [@Zahavi1975]. This is possible whenever the benefits of signaling outweigh the costs for those who have the relevant underlying quality but the costs outweigh the benefits for those who lack the quality or possess it to a lesser degree [@zahavi1977; @grafen1990]. In particular, if the benefits of sending the signal are the same for signalers regardless of condition, but the signal cost outweighs this benefit for signalers without the underlying quality (a "costly signal"), then individuals with the underlying quality can evolve to send the signal and receivers can evolve to respond in mutually beneficial ways to the signal.

@@

Beyond signaling need to those invested in the well-being of the signaler, depression and suicidal behavior may also help an individual bargain for increased support [@Hagen2003]. This is possible due to the reduced investment in others that frequently accompanies depression or the threat to future cooperation that comes from suicide. In these instances, depression and suicide attempts may function analogously to a labor strike, sending the message that one’s need is genuine and that if the target of the signal wanted to continue receiving benefits from the depressed, they would need to increase support towards the signaler. Due to this bargaining component, aversive responses from the targets of the signal to the signaler are expected throughout the bargaining process, with potential helpers likely signaling the costliness of providing support. Importantly, such responses do not prevent the delivery of meaningful benefits, they are simply expected to reduce the risk of helpers providing too much support for too little need.

<!-- reword this paragraph -->

@@

## Responses to depression and suicidal behavior

For depression and suicide attempts to be worth the costs involved with signal creation, meaningful social support would be required [@Hagen2003; @SymeEtAl2016]. Most studies examining responses to depression have been observational and focused on responses outside of changes in social support. In general, most report aversive responses and high levels of rejection towards the depressed, and depressed individuals are often linked to impaired social functioning [@EvraireDozois2011; @GotlibLee1989; @HirschfeldEtAl2000]. One explanation for this is that depressed individuals engage in behaviors that drive others away. In line with this reasoning, both excessive reassurance seeking and negative feedback seeking, which often accompany depression, have been shown to elicit negative emotional responses from others and increase the chance of rejection from others [@JoinerMetalsky1995; @StarrDavila2008]. Negative interactions are also seen in families as depressed children tend to report less support and more negative responses from parents than non-depressed children, although it has not been demonstrated that such interactions did not lead to depression in the first place [@MesserGross1995; @SchwartzEtAl2012].

Despite being looked for less frequently, beneficial responses to depressed individuals are also reported. Depression has been associated with reduced likelihood of aggression within families [@hops1987; @dadds1992], including those already experiencing high levels of conflict [@SheeberEtAl2001]. There is also limited evidence that depression can result in increased levels of support from others outside the family. For example, non-depressed individuals were shown to be more cooperative and exhibit more caretaking behavior towards their college depressed roommates over the course of their time living together [@HokansonEtAl1986]. Likewise, a study examining responses to confederates trained to appear ‘depressed’, ‘anxious’, or ‘normal’ found that individuals were more likely to provide advice and support to ‘depressed’ individuals than ‘anxious’ or ‘normal’ individuals despite the confederate in the ‘depression’ condition being viewed more negatively than the other conditions and the ‘depression’ condition involving similar levels of distress compared to the ‘anxious’ condition [@stephens1987].

Most studies examining responses to those who have survived suicide attempts have focused on stigmatization. In these studies, perceptions of survivors as being weak, selfish, mentally ill, and antisocial are all being commonly reported [@BatterhamEtAl2013; @TzengLipson2004], with stigmatization often being found within and outside one's social network [@FreyEtAl2016; @ScoccoEtAl2012]. Despite this potential for stigmatization, increased social support and changes to important relationships have been reported to follow suicide attempts with some indication that these effects may hold long term [@Stengel1956]. For example, a study of 100 women who survived suicide attempts found that individuals gained identifiable benefits through the attempt in 75 cases, with 41 individuals benefiting from reconciliations with others [@Lukianowicz1971]. Unlike many Western countries, where suicide is often viewed as pathological [@Hidaka2012], members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness [@Shostak1981], with attempts having been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers [@GutierrezdePinedaMuirden1948; @Hilger1957; @Karsten1935; @Tessmann1930; @Wilson1960]. Although findings that individuals view suicide attempts as cries for help is not necessarily evidence for their ability to lead to beneficial responses, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, despite suicide rarely if ever being the main topic [@SymeEtAl2016].

@@

Most of the limited literature on social responses to depression and suicidality comprises observational studies of depressed individuals interacting with family, friends, or roommates. These have ecological validity, but cannot easily determine causal relationships. Some studies, though, have employed an experimental design in which participants were randomized into conditions in which they listened to, watched, or interacted with either a depressed or non-depressed person, where in some cases the depressed person was a non-depressed confederate enacting a depressed role [@marcus1992]. These designs can demonstrate causation but the transient, inconsequential relationships and laboratory settings lack ecological validity.

@@

The "romantic partner" vignette had participants imagine that they found a highly desirable partner after years of being single. However, their 13-year-old daughter, who has a history of trying to get in the way of their relationships is clearly not happy with this. After a period of sustained conflict, she accuses the romantic partner of physically assaulting her but cannot produce any evidence.

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is described as trying to arrange a dowry for their older daughter so she can marry a man she already loves. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. Due to the younger daughter in the story only being two years younger, any additional investment in the dowry will come at the younger daughter's expense.

@@

We used the pilot data to estimate the sample sizes needed to detect an effect of the *Depression* signal vs. *Verbal request* control (two conditions) on *Belief* in the victim's need. Power was about 80% for a sample size of about `r pwr80_90[[2,2]]`, and was about 90% for a sample size of about `r pwr80_90[[2,3]]`. We also estimated the sample size needed to detect an effect of *Depression* vs. a *Control* condition that was meant to not include any signaling (but which we realized, post hoc, did include a verbal request). Power was about 80% for a sample size of about `r pwr80_90[[1,2]]`, and was about 90% for a sample size of about `r pwr80_90[[1,3]]`. See Figure \@ref(fig:power). Given our $1500 USD budget, we aimed for a sample size of 125 for two conditions, and 1250 for all conditions in the study. For more details, see the SI and XX.

>>>>

Major depression and suicide are generally viewed as pathological responses to stress that elicit aversive responses from others. Alternatively, depression and suicide attempts might involve honest 
signals of need in response to stress and help to bargain for increased support from reluctant 
others. To address this hypothesis, we examined responses to emotional signals in an experimental vignette study involving claims of substantial need in the presence of conflicts of interest and private information about whether the signaler was actually in need. In our sample of 1,240 participants recruited from Amazon Mechanical Turk, costlier signals like depression and suicide attempts resulted in greater perceptions of need and increased likelihood of support compared to simple verbal requests and crying without further symptoms. Additionally, the effect of signal type on likelihood of support was largely mediated by the effect of signaling on belief in need as predicted. Our results suggest that depression and suicide attempts, apparent human universals, are relatively successful ways of signaling honest need and eliciting support, consistent with costly signaling explanations.

@@

Durkheim’s pioneering study of suicide [@durkheim1897] helped launch sociology as a quantitative discipline. Sociological investigations of suicide flourished through much of the 20th century, but were supplanted by medical and public health approaches as the biomedical model of mental health took hold in the late 1970's and early 1980's [@wray2011; @syme2020]. The biomedical model tends to frame suicide as a problem with the individual, e.g., as one involving psychiatric disturbance, drug addiction, or stress sensitivity [@wray2011]. This theoretical approach has made little progress: rates of suicide, which vary widely by country, have remained relatively flat in countries for which there are long-term data, such as the US, UK, Japan, and Sweden [@owidsuicide].

@@

Evolutionary theories of depression tend to either ignore suicidality, or take it as evidence that depression is dysfunctional. Wakefield, for instance, who has devoted much of his career to arguing that most cases of depression are simply severe but normal sadness and grief, nevertheless takes suicidality to indicate pathology for some cases of depression [e.g., @wakefield2017]. @nettle2004 argues that the costs of depression, including suicidality, outweigh any benefits, and therefore adaptationist models of depression are implausible. @gilbert2006 sees suicidality as an example of a dysregulated escape defense [for reviews of evolutionary theories of depression, see @Hagen2011].

<!-- and anthropological explanations for low mood have become more common. In general, these approaches see sadness and minor depression as ways to motivate adaptive behavior changes in response to negative outcomes in one's environment (cite Nesse and others), mainly differing in how they view major depression. While some approaches present most forms of major depression as dysfunctional (inflammation and some of Nesse), others have extended the logic on the benefits of low mood and reduced behavior to more severe instances of sadness. Of the possible benefits, the ability of depression to help solve fitness-relevant in times of social conflict is the most emphasized, with depression being suggested to help by reallocating resources with the body for problem solving (cite ARH), reducing the costs of interacting with higher status individuals (cite), or serving as a means to both signal one's need and bargain for increased support (cite).  -->

<!-- As with suicide, research on depression since the 1980's has been dominated by biomedical approaches of limited success [@patten2016; @ormel2019].  -->

<!-- and there is still substantial uncertainty about why both are as common as they are (cite). Likely due to  the suffering they cause, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others). -->

@@

<!-- Likely the most reliable finding regarding depression and suicidality is that most cases are closely tied to major negative life events [@Hammen2005; @Mazure1998; @devries2011; @ellsberg2008]. Compared to non-depressed individuals, those with depression tend to report at least twice as many instances of negative events than non-depressed individuals [@Mazure1998] and more negative events than those with schizophrenia and bipolar depression across multiple studies [@Paykel1994]. In longitudinal studies, the onset of depression has been reported to quickly follow the negative event [@KendlerEtAl1999] or coincide with periods where conflict is likely to increase prior to it (e.g., depression starting before a divorce rather than after; @rosenstrom2017).  -->
<!-- <!-- look for other longitudinal studies --> -->

<!-- This strong connection between negative life events and depression along with the adversity and distress that likely accompanies these events has led to the suggestion that this relationship is causal [@KendlerEtAl1999]. Support for this comes from the fact that the relationship between negative events and depression holds true when looking solely at events outside of one's control, showing that the connection is unlikely to be driven solely by individuals who are already depressed creating situations where negative events are likely to be common [@Hammen2005; @KendlerEtAl1999]. In addition to this, twin studies have shown that one's history of negative events is still a strong predictor of major depression when controlling for genetic similarity and that part of the heritability of depression stems from the heritability of negative events, like divorce and family conflict [@KendlerBaker2007; @KendlerEtAl1999]. -->

Likely the most reliable finding regarding depression and suicidality is that most cases are closely tied to major negative life events [@Hammen2005; @Mazure1998; @devries2011; @ellsberg2008], with such events likely causing depressive states based on twin studies and examinations involving negative events outside of one's control[@Hammen2005; @KendlerEtAl1999]. Of the types of negative life events associated with depression, those which involve conflict tend to be more strongly associated with depression [@Hammen2005; @Mazure1998]. Marital problems, bullying, and the existence of abusive relationships are all commonly cited risk factors for depression [@KendlerEtAl1995; @KendlerEtAl1999; @klomek2007; @klomek2019], with  assault (both sexual and non-sexual), in particular, greatly increasing one's risk of depression [@KendlerEtAl1995; @KendlerEtAl1999].  Other notable risk factors like loss of a loved one, financial problems, job loss, or severe or prolonged illness, might seem less related to conflict. However, the ability for others to provide meaningful support to limit the harm caused in times of need [@SugiyamaSugiyama2003] combined with the ubiquity of conflicts of interests mean that many types of problems will eventually involve social conflict even if they did not start that way [@Hagen2003]. In line with this reasoning, conflict, as a whole, was found to be associated with depression among the Tsimane, particularly when it involved non-kin [@StieglitzEtAl2015a].

Due to depression's relationship to sucide [@who2001; @hawton2013], it should not be surprising that suicidality also often coincides with social adversity and conflict. Grief over the loss of a loved one [@firth1936: @Firth:1961; @devries2011], bullying [@alavi2017; @butler2019; @ford2017; @wang2020], and instances of social transgression [@Firth:1961; @Malinowski:1932; @SymeHagen2018; @lester1997] have all been reported to lead to suicidal behavior in both industrialized and traditional societies. Of the recorded instances of suicide attempts in the Human Relations Area Files (HRAF), 77.4% of attempts were found to be linked to times of social conflict, 88.7% were linked to fitness threats in general, and 84.9% were linked to situations in which attempters were powerless to resolve their situation alone [@SymeEtAl2016]. As with depression, assault is an particularly powerful risk factor for suicidal behavior based on multiple large-scale multi-country studies [@devries2011; @ellsberg2008] and more localized examinations [@reigstad2017].

## Depression and suicide attempts as both signals and bargaining tools

@@

Due to the ease with which signals can be faked [@Zahavi1975; @mokkonen2016] and the costs of accepting a deceptive signal as reliable [@dawkins1978; @johnstone1993; @Hasson1994], demonstrating that one’s need is genuine has likely been a recurrent adaptive problem since humans evolved the capacity to provide instrumental support to one another. One potential solution to this is to produce signals that are unprofitable to fake [@Zahavi1975]. This is possible whenever the benefits of signaling outweigh the costs for those who have the relevant underlying quality but the costs outweigh the benefits for those who lack the quality or possess it to a lesser degree [@zahavi1977; @grafen1990]. In particular, if the benefits of sending the signal are the same for signalers regardless of condition, but the signal cost outweighs this benefit for signalers without the underlying quality (a "costly signal"), then individuals with the underlying quality can evolve to send the signal and receivers can evolve to respond in mutually beneficial ways to the signal.

@@

<!-- Beyond signaling need to those invested in the well-being of the signaler, depression and suicidal behavior may also help an individual bargain for increased support [@Hagen2003]. This is possible due to the reduced investment in others that frequently accompanies depression or the threat to future cooperation that comes from suicide. In these instances, depression and suicide attempts may function analogously to a labor strike, sending the message that one’s need is genuine and that if the target of the signal wanted to continue receiving benefits from the depressed, they would need to increase support towards the signaler. Due to this bargaining component, aversive responses from the targets of the signal to the signaler are expected throughout the bargaining process, with potential helpers likely signaling the costliness of providing support. Importantly, such responses do not prevent the delivery of meaningful benefits, they are simply expected to reduce the risk of helpers providing too much support for too little need. -->

@@

<!-- ## Responses to depression and suicidal behavior  -->

<!-- For depression and suicide attempts to be worth the costs involved with signal creation, meaningful social support would be required [@Hagen2003; @SymeEtAl2016]. Most studies examining responses to depression have been observational and focused on responses outside of changes in social support. In general, most report aversive responses and high levels of rejection towards the depressed, and depressed individuals are often linked to impaired social functioning [@EvraireDozois2011; @GotlibLee1989; @HirschfeldEtAl2000]. One explanation for this is that depressed individuals engage in behaviors that drive others away. In line with this reasoning, both excessive reassurance seeking and negative feedback seeking, which often accompany depression, have been shown to elicit negative emotional responses from others and increase the chance of rejection from others [@JoinerMetalsky1995; @StarrDavila2008]. Negative interactions are also seen in families as depressed children tend to report less support and more negative responses from parents than non-depressed children, although it has not been demonstrated that such interactions did not lead to depression in the first place [@MesserGross1995; @SchwartzEtAl2012]. -->

<!-- Despite being looked for less frequently, beneficial responses to depressed individuals are also reported. Depression has been associated with reduced likelihood of aggression within families [@hops1987; @dadds1992], including those already experiencing high levels of conflict [@SheeberEtAl2001]. There is also limited evidence that depression can result in increased levels of support from others outside the family. For example, non-depressed individuals were shown to be more cooperative and exhibit more caretaking behavior towards their college depressed roommates over the course of their time living together [@HokansonEtAl1986]. Likewise, a study examining responses to confederates trained to appear ‘depressed’, ‘anxious’, or ‘normal’ found that individuals were more likely to provide advice and support to ‘depressed’ individuals than ‘anxious’ or ‘normal’ individuals despite the confederate in the ‘depression’ condition being viewed more negatively than the other conditions and the ‘depression’ condition involving similar levels of distress compared to the ‘anxious’ condition [@stephens1987].  -->

<!-- Most studies examining responses to those who have survived suicide attempts have focused on stigmatization. In these studies, perceptions of survivors as being weak, selfish, mentally ill, and antisocial are all being commonly reported [@BatterhamEtAl2013; @TzengLipson2004], with stigmatization often being found within and outside one's social network [@FreyEtAl2016; @ScoccoEtAl2012]. Despite this potential for stigmatization, increased social support and changes to important relationships have been reported to follow suicide attempts with some indication that these effects may hold long term [@Stengel1956]. For example, a study of 100 women who survived suicide attempts found that individuals gained identifiable benefits through the attempt in 75 cases, with 41 individuals benefiting from reconciliations with others [@Lukianowicz1971]. Unlike many Western countries, where suicide is often viewed as pathological [@Hidaka2012], members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness [@Shostak1981], with attempts having been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers [@GutierrezdePinedaMuirden1948; @Hilger1957; @Karsten1935; @Tessmann1930; @Wilson1960]. Although findings that individuals view suicide attempts as cries for help is not necessarily evidence for their ability to lead to beneficial responses, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, despite suicide rarely if ever being the main topic [@SymeEtAl2016]. -->

@@

Most of the limited literature on social responses to depression and suicidality comprises observational studies of depressed individuals interacting with family, friends, or roommates [@hops1987; @dadds1992; @SheeberEtAl2001; @JoinerMetalsky1995; @StarrDavila2008]. These have ecological validity, but cannot easily determine causal relationships. Some studies, though, have employed an experimental design in which participants were randomized into conditions in which they listened to, watched, or interacted with either a depressed or non-depressed person, where in some cases the depressed person was a non-depressed confederate enacting a depressed role [@marcus1992]. These designs can demonstrate causation but the transient, inconsequential relationships and laboratory settings lack ecological validity.

@@

The "romantic partner" vignette had participants imagine that they found a highly desirable partner after years of being single. However, their 13-year-old daughter, who has a history of trying to get in the way of their relationships, is clearly not happy with this. After a period of sustained conflict, she accuses the romantic partner of physically assaulting her but cannot produce any evidence.

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family are described as trying to arrange a dowry for their older daughter so she can marry a man she already loves. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. Due to the younger daughter in the story only being two years younger, any additional investment in the dowry will come at the younger daughter's expense.

@@

We used the pilot data to estimate the sample sizes needed to detect an effect of the *Depression* signal vs. *Verbal request* control (two conditions) on *Belief* in the victim's need. Power was about 80% for a sample size of about `r pwr80_90[[2,2]]`, and was about 90% for a sample size of about `r pwr80_90[[2,3]]`. We also estimated the sample size needed to detect an effect of *Depression* vs. a *Control* condition that was meant to not include any signaling (but which we realized, post hoc, did include a verbal request). Power was about 80% for a sample size of about `r pwr80_90[[1,2]]`, and was about 90% for a sample size of about `r pwr80_90[[1,3]]`. See Figure \@ref(fig:power). Given our $1500 USD budget, we aimed for a sample size of 125 for two conditions, and 1250 for all conditions in the study. For more details, see the SI and /.
